The symmetry of partner modelling

### 2016-02-04 mutual-modeling.tex
Updated section 'Synthesis'

Out of 5 studies, 3 revealed a correlation between the model peers built about
each other: when $A$ builds an accurate model of $B$, $B$ also tends to build an
accurate model of A.  In one study, this was the case only in one condition.
The conclusion is hence that modelling is mutual but only in certain conditions,
summarized in Table~\ref{synthesis_table}.

\begin{table*}[ht!]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{p{2.6cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}}
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
\textbf{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
\textbf{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
\textbf{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
\textbf{Group Size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\textbf{Awareness tool}       & Partner's position               & Partner's current/past pos.     & -                         & Partner's concept map     & Partner's scores at pre-test \\
\textbf{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge       & Team/Ind. knowledge          \\
\textbf{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
$\mathcal{H}_2$               &                                  & supported&                         &                           &                              \\
$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
\end{tabular}
}
\caption{Comparison of the studies}
\label{synthesis_table}
\end{table*}

Coming back to our initial hypotheses, the $\Delta_2$ results in study 2
reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The \emph{interaction
hypothesis} $\mathcal{H}_{3}$ was that the activity of modelling one's partner
is not simply the result of an individual cognitive effort but overall the
side-effect of the quality of interactions. The \emph{second-order-modelling
hypothesis} $\mathcal{H}_{2}$ stated that, if $A$ is accurate in modelling $B$,
$A$ should notice when $B$ misunderstands $A$ and should be able to repair. A
good modeler would somehow drive the modelling of his or her partner.  This
\emph{drive} will however occur through verbal cues and repairs and is hence not
easy to differentiate from $\mathcal{H}_{3}$. The contradictory $\Delta_2$
analysis in study 3 and $\Delta_1$ results in studies 4 (to some extent) and 5
weaken the hypothesis.

The triangular questions lead to similar contradictions. The $\Delta_2$
correlation between the \gmodel{A}{C} and \gmodel{B}{C} (study 2)
support $\mathcal{H}_{2}$ if modelling skills were purely individual, there would
be no reason for modelling skills of $A$ and $B$ to be correlated.  The role of $C$ in
\gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (study
3) question $\mathcal{H}_{2}$: if the role of the modellee was so
important, it would be surprising to have by chance a correlation between
$B$'s and $C$' ability to drive $A$'s modelling.

In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
($\Delta_4$ in study 4 and study 5) could also contribute to the idea that $A$'s
modelling skills is not the key factor. However, between modelling someone and
modelling oneself, \ie  metacognition, they are so many differences, that we
would hardly conclude from that.

Among these complex results, it is important to stress that, for $\Delta_2$, the
significant correlations are all below 0.50, not around 0.90.  So even if all
$\Delta_2$ correlations were significant in all experiments, we would still not
conclude that modelling purely reflects the quality of social interactions.
There is obviously a large part of individual variance; some of us are much
better in modelling peers. This is not new. The contribution of these studies is
rather to establish the importance of team processes on something that a priori
seemed to be more of a personal cognitive skill.  In other words, we stress that
the bottle is half-full.

These conclusions must be presented with multiple disclaimers. First, they
heavily rely on correlations; hence we cannot identify causal links. Second, we
faced hard methodological issues. Providing learners with on-task questionnaires
introduce a bias: they will pay more attention to their partners in the
remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
and rationalization biases. The nature of mutual modelling implies methodological
challenges that call for new  measurement methods. We have promising results for
using eye tracking methods to address this challenge.

Finally, let us stress the expressions we used, such as \gmodel{A}{B}, do not
imply we have a mechanical view of modelling. It's just a language for referring
to different components of modelling each other in teams.  This simple notation
emerges from the need to compare five mutual modelling studies; it did not
generate a consistent program of consistent studies.  Our results are not very
robust because they emerge from a \emph{post-hoc} reinterpretation of studies
that addressed different research questions (media richness, awareness tools,
scripts,...). This diversity makes our results difficult to integrate, partly
contradictory. Nonetheless, this diversity gives some generalizability: mutual
modelling has been investigated in different contexts (virtual space versus real
space), with different groups sizes (pairs and triads) and different tasks.
Moreover, these results have been found using different methods for measuring
modelling: on-task in study 1 versus off-task in study 2, subjective validation
in study 1 versus objective validation in study 2 (comparing $A$'s model with $B$'s
behavior).

>>>>

We found symmetry of the mutual model on 4 studies out of 5. In study 4, this only applies to the control group (having read the same text before), which corresponds to the situation of the 3 first studies.  Finding this symmetry per se constitutes an interesting result as we don't know earlier studies who have established this relationship. The problem is that this symmetry does not allow use to discriminate  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The second hypothesis, the skill of helping one's peer to model oneself, is questioned by study 3 where C could be modelled  accurately by A and not accurately by B. The same hypothesis is however supported by the results of study 2, and indirectly by study 4.  This does not discard \mathcal{H}_{1}$: even when we found significant correlations, they were all below 0.50, not around 0.90.  Hence, even of the quality of social interaction matters, there is obviously a large part of individual variance within teams.

In other words, these studies do conclude that one hypothesis is right and the others are wrong, which is per se the main contribution of this paper. We started from the idea that partner modelling is essentially an individual skill  and that we therefore improve the quality of collaboration by providing the so-called awareness tools, which act as some kind of prostheses for partner modelling. This individual skill cannot be discarded.  Everyone has experienced the pleasure of interacting with colleagues who answer precisely to the question we asked them  and even guess the reason why we asked this question. Conversely, everyone also experienced the frustration of someone referring about a third person by his name, say Mike Smith,  while knowing perfectly that there is no chance that we know this Mike Smith. The existence of these individual skills is not denied by our studies; the novelty of this paper is to show, thanks to symmetry values, that individual skills only account  for a part of the accuracy in mutual modelling.

We acknowledge that the difference between these hypotheses is  rather theoretical since the  process of modelling one's partner( $\mathcal{H}_{1}$) and the process of helping one's partner to model onself ($\mathcal{H}_{3}$) are mediated by verbal interactions in the team. One could hardly imagine someone managing a very accurate modelling despite the very bad quality of interactions.  There is a bidirectional causal link between the accuracy of mutual modelling and the quality of interactions. This rather artificial distinction does however participate to the more fundamental discussion of the role of individual and social mechanisms in human cognition. In the field of social cognition, it is a common place to state that 'the whole is greater than the sum of the parts'. This refers to the emergence of team properties than cannot be reduced to the set of individual contributions.  Our paper illustrates this emergence by showing the symmetry of mutual modelling

These conclusions must be presented with multiple disclaimers. First, they
heavily rely on correlations; hence we cannot identify causal links. Second, we
faced hard methodological issues. Providing learners with on-task questionnaires
introduce a bias: they will pay more attention to their partners in the
remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
and rationalization biases. The nature of mutual modelling implies methodological
challenges that call for new  measurement methods. We have promising results for
using eye tracking methods to address this challenge. 
Third, our results emerge from a \emph{post-hoc} reinterpretation of studies
that addressed different research questions (media richness, awareness tools,
scripts,...). This diversity makes our results difficult to integrate, partly
contradictory. Nonetheless, this diversity gives some generalizability: mutual
modelling has been investigated in different contexts (virtual space versus real
space), with different groups sizes (pairs and triads) and different tasks.

Finally, let us repeat that the expressions we used, such as \gmodel{A}{B}, do not
imply we have a mechanical view of modelling. It's just a language for referring
to different components of modelling each other in teams.  This simple notation
emerged from the need to compare five mutual modelling studies. It could however pave the road for elaborating a systematic agenda for research on mutual modelling.
